---
title: "EDA"
format: html
editor: visual
---

## Diabetic Data Introduction

```{r}
main<- read.csv("C:/Users/James P/Downloads/diabetes_binary_health_indicators_BRFSS2015.csv")
head(main)
```

Above is the head of the data set we are working with. This is a diabetes health indicator set. We will be using the binary Diabetes variable as our response. To note, the site from which the data set originates, [kaggle](https://www.kaggle.com/datasets/alexteboul/diabetes-health-indicators-dataset/), fails to mention whether this is type 1 or type 2 diabetes. The site also mentions that diabetes increases risk of stroke, heart attack, and high blood pressure, so these variables will not be considered as predictors.

```{r, echo = F}
library(caret)
library(dplyr) #Need to clean this up a bit
main[,-5]<- lapply(main[,-5], factor)
```

## Analysis

Firstly, for the predictors with two levels, I will borrow the confusion matrix from the caret package to check how well they align for the modelling.

```{r}
confusionMatrix(main$Diabetes_binary, main$HvyAlcoholConsump)
```

We can see for heavy alcohol consumption, with a 1 for consumption and 0 for non-heavy consumption, that there is an 85% accuracy in predicting diabetes.

```{r}
confusionMatrix(main$Diabetes_binary, main$Veggies)
```

For veggies, we can see that most people who eat veggies do not have diabetes. This negative information could be useful for the model.

```{r}
library(ggplot2)
ggplot(main, aes(x = BMI, color = Diabetes_binary)) + geom_bar()
```

BMI, or Body Mass Index, may not be a great indicator of diabetes, but the tail certainly seems to be more blue than orange.

```{r}
ggplot(main, aes(x = Age, color = Diabetes_binary)) + geom_bar()
```

Diabetes appears to occur in the oldest age groups.

## Discussion

For the modelling page, I am going to try to fit models using BMI, Age, Veggies, and Alcohol as predictors for the the response. Since alcohol already fit diabetes so well, hopefully the other three will improve the fit. Since type 2 diabetes takes time to manifest, age should assist in that prediction. Lastly, someone who consumes vegetables will be at less risk, so there will be a "negative" term in the model.
